This thesis investigates discretization robustness in Fourier- and convolution-based neural operators on both uniform Cartesian grids and non-uniform point-cloud domains. On uniform grids, the Fourier Neural Operator (FNO) and the Convolutional Neural Operator (CNO) are evaluated on Navier--Stokes and Darcy flow benchmarks through resolution-transfer experiments. The results show that CNO exhibits substantially stronger robustness across resolutions, while FNO is more sensitive to discretization mismatch due to the resolution dependence of its retained Fourier modes. An ablation study shows that the dominant source of this instability is not nonlinear aliasing, but the lack of a fixed internal representation; introducing resampling onto a fixed internal grid reduces the discretization sensitivity of FNO by nearly two orders of magnitude.

For non-uniform meshes, the geometry-aware Fourier Neural Operator (Geo-FNO) is evaluated on incompressible flow past a cylinder and on a reactor-relevant high-temperature gas-cooled reactor (HTGR) temperature dataset. The results show that Geo-FNO can learn meaningful predictions directly on irregular point clouds and non-uniform meshes. At the same time, discretization invariance in the mesh-free setting must be interpreted differently than on structured grids, as robustness to changing sampling geometry rather than resolution transfer. A reconstruction bug in the original Geo-FNO implementation was also identified and corrected, removing a persistent seam artifact and significantly improving predictive accuracy.

Overall, the results show that discretization robustness depends strongly on how neural operators control the relationship between continuous functions and their discrete representations. Stable generalization requires explicit control of bandlimiting, internal resolution, and representation consistency across discretizations.
