\chapter{\label{cha:background}Background \& Theory}

This chapter introduces the theoretical background needed for the
remainder of the thesis. It first places operator learning in context
with more classical learning setups and then discusses the neural
operator architectures that are central to this work.

\section{Operator Learning}

This section introduces operator learning as a mapping between
functions, rather than the more familiar vector-to-vector setting. It
should explain why this perspective is useful for PDE-governed physical
systems and introduce the CDE concept that motivates resolution-aware
learning.

\subsection{Vector-to-vector versus function-to-function mapping}

Contrast conventional supervised learning with operator learning and
explain why PDE solution operators are naturally formulated in the
function-to-function setting.

\subsection{Continuous-domain perspective}

Introduce the continuous-domain view and connect it to the notion of
discretization invariance used throughout the thesis.

\section{Fourier Neural Operator}

This section should explain the FNO architecture and the spectral
convolution mechanism on which it is based.

\subsection{Architecture}

Describe the lifting layer, Fourier layers, spectral convolution,
non-linear transformations, and projection back to the target field.

\subsection{Strengths}

Discuss the global receptive field, computational efficiency through
FFT-based convolutions, and the suitability of FNO for smooth
solution operators.

\subsection{Weaknesses}

Discuss spectral truncation, sensitivity to resolution changes, and
the reliance on regular grids.

\section{Convolutional Neural Operator}

This section should explain the CNO architecture and the inductive
biases that differentiate it from FNO.

\subsection{Architecture}

Describe the encoder-decoder style structure, convolutional blocks, and
the way multi-scale processing is used.

\subsection{Bandlimited assumption}

Explain the bandlimited assumption and why it can stabilize learning and
support more consistent behavior across discretizations. Discuss how this means explicitly discarding high-frequency information, where the FNO keeps it (but does not know what to do with it)

\subsection{Local versus global inductive bias}

Compare the locality of convolutional operators with the global spectral
bias of FNO and connect this to the later experiments.

\section{Mesh-free Operator Learning and geo-FNO}

This section motivates operator learning on irregular domains and
introduces geo-FNO as the architecture used for this purpose.

\subsection{Motivation for irregular domains}

Explain why many scientific and engineering problems do not naturally
live on uniform Cartesian grids and why this matters for neural
operators.

\subsection{geo-FNO}

Describe the geo-FNO approach and how it extends Fourier-based operator
learning to non-uniform meshes.

\section{Discretization invariance on mesh-free grids}

More philosophical discussion of what discretization invariance means when we no longer have a regular grid and how this relates to the notion of a continuous-domain operator. 
