\chapter{\label{cha:conclusion}Conclusions and Future Work}

\section{Conclusions}

The goal of this thesis was to investigate to what extent Fourier- and convolution-based neural operators can provide discretization-robust surrogate models for PDE-based data on both uniform grids and non-uniform point clouds. Particular focus was placed on discretization invariance: whether a model learns the underlying continuous operator or becomes tied to the specific discretization used during training.

For uniform Cartesian grids, the results show a clear difference between the FNO and the CNO. While the FNO achieves strong accuracy at the training resolution, its performance degrades significantly under resolution transfer, especially when evaluated on coarser grids. This behavior is caused primarily by the dependence of retained Fourier coefficients on the training discretization. As the evaluation resolution changes, the physical meaning and energy distribution of the retained modes change, causing the learned correspondence between Fourier coefficients and physical structure to break down.

The comparison between FNO\_k8 and FNO\_k16 further highlights this trade-off. Retaining more modes improves expressivity and lowers the in-distribution loss, but also increases sensitivity to discretization mismatch. The stronger performance of FNO\_k16 at all tested resolutions shows that additional spectral information is useful, but its much larger increase in loss away from the training resolution demonstrates that this improved accuracy comes at the cost of robustness.

In contrast, the CNO exhibits substantially stronger discretization robustness across both Navier--Stokes and Darcy flow. Its use of fixed internal resampling provides a stable representation across resolutions, while the multi-scale encoder--decoder structure allows the model to capture both local and global behavior without relying on resolution-dependent Fourier modes. The results show that this architectural choice is more important for robustness than the specific use of spectral representations.

The role of the upsample--nonlinearity--downsample (UND) block was found to be more subtle. The UND block improves mathematical consistency by preventing nonlinear aliasing from contaminating the retained modes, but it is not the dominant factor controlling discretization robustness. In the modified FNO experiments, introducing UND alone produced only a negligible improvement, while resampling onto a fixed internal grid reduced discretization sensitivity by nearly two orders of magnitude. This shows that representation consistency is the primary mechanism governing robustness, while aliasing control is a secondary but still theoretically important refinement.

For non-uniform meshes, the results show that the Geo-FNO successfully extends operator learning to irregular spatial domains. On both the cylinder flow dataset and the HTGR temperature dataset, the model produced physically meaningful predictions directly on point clouds and non-uniform meshes without requiring interpolation onto a uniform Cartesian grid. This demonstrates that Fourier-based operator learning can remain effective even when the assumptions of the standard FNO no longer hold.

At the same time, the thesis shows that discretization invariance must be interpreted differently in the mesh-free setting. On uniform grids, robustness can be evaluated through resolution transfer by changing grid spacing while preserving the same domain. For point clouds, there is no single notion of resolution; the quality of the representation depends not only on the number of points, but also on where those points are placed relative to important physical features. As a result, mesh-free discretization invariance is better understood as robustness to changing sampling geometry rather than transfer across resolutions.

The cylinder flow results also highlight an important practical limitation of one-shot operator learning for unsteady problems. The largest prediction errors occurred in the downstream wake, where some samples exhibited oscillatory behavior associated with transient wake dynamics or vortex shedding. In these cases, Geo-FNO accurately predicted the mean flow structure but often smoothed the instantaneous wake phase. This indicates that part of the remaining error is caused by dataset structure and problem formulation rather than by a failure of the model to learn the dominant physics.

Overall, the results of this thesis show that discretization robustness depends primarily on how a neural operator handles the relationship between continuous functions and their sampled representations. Models that enforce a stable internal representation, such as the CNO or FNO variants with fixed internal resampling, generalize more reliably across discretizations than models that rely directly on resolution-dependent spectral coefficients. For irregular domains, the same principle remains central: reliable operator learning depends not only on model capacity, but on how consistently the architecture represents the underlying continuous problem across different discretizations.   




\section{Future Work}

Several natural extensions of this work remain for future study.

A first priority is the use of more challenging non-uniform reactor datasets. While the HTGR temperature results demonstrate that Geo-FNO can be applied to reactor-relevant data on irregular meshes, the problem itself is relatively simple: the geometry is fixed across all samples, the input consists only of boundary temperature and axial position, and the resulting temperature field is smooth and strongly structured. This makes the experiment closer to a proof of applicability than a demonstration of performance on realistic reactor simulations. Future work should therefore consider datasets involving stronger nonlinear coupling and more complex physical behavior, such as transient thermal-hydraulics, coupled neutronics and heat transfer, spatially varying material properties, or changing geometries caused by fuel movement or control rod insertion. Such datasets would provide a more meaningful test of whether neural operators can function as practical surrogate models in reactor analysis.

A second important extension is a more systematic study of discretization invariance on non-uniform meshes. In this thesis, no direct resolution-transfer experiments were performed for mesh-free datasets because there is no single notion of resolution for point clouds. Prediction quality depends not only on the number of points, but also on where those points are placed relative to boundaries, sharp gradients, and other important physical features. Future work should therefore construct controlled families of point clouds where local density is varied in a physically meaningful way, for example by refining sampling near cylinder boundaries, wake regions, or thermal boundary layers while keeping the global geometry fixed. This would make it possible to evaluate robustness to sampling geometry directly rather than relying only on held-out prediction quality.

Another limitation of the mesh-free experiments is that only one architecture, Geo-FNO, was evaluated. For the uniform-grid experiments, the contrast between the globally biased FNO and the more locally biased CNO provided an important comparison. For non-uniform meshes, no equivalent comparison was made. This is particularly relevant for the cylinder flow dataset, where the dominant dynamics are strongly local and a local architecture may perform better than Geo-FNO. Future work should therefore compare Geo-FNO against graph-based neural operators or other locally structured mesh-free models, as well as against the common baseline of interpolating the data onto a uniform Cartesian grid and applying a standard FNO. Such comparisons would help determine whether the advantages of Geo-FNO come primarily from learning directly on the original point cloud or from the spectral structure of the architecture itself.

A particularly interesting architectural extension would be the development of a geometry-aware Convolutional Neural Operator (Geo-CNO). One possible approach would be to combine the learned coordinate transformation 
$\phi^{-1}$ from Geo-FNO with the CNO architecture by first mapping the irregular point cloud to a more regular computational representation and then applying a CNO instead of an FNO. This would preserve the local inductive bias of convolutional operators while also benefiting from fixed internal resampling and explicit control of bandlimiting. Such an architecture could help determine whether the strong discretization robustness of CNO on uniform grids can be extended to irregular domains and whether locality can improve performance on datasets such as cylinder flow, where Geo-FNO tends to struggle most in the wake region. This direction was also suggested in the original CNO paper and would provide a natural continuation of the work in this thesis.

The cylinder flow results also suggest that future work should investigate arbitrary-time prediction rather than only one-shot prediction of a fixed final state. The largest prediction errors occurred in samples with oscillatory wake behavior, where Geo-FNO often predicted the correct mean wake location but smoothed out the instantaneous downstream oscillations. This indicates that the model captures the dominant flow structure but struggles to predict the exact phase of the unsteady wake. A natural extension would therefore be autoregressive rollout, where the model predicts short time increments and repeatedly advances the solution forward in time rather than predicting the distant final state directly. Techniques such as teacher forcing, recurrent rollout stabilization, or hybrid one-shot and autoregressive schemes could make these predictions more stable. This would likely be particularly beneficial for transient fluid problems where phase accuracy matters more than prediction of the mean flow alone.

For uniform-grid neural operators, the modified FNO experiments show that fixed internal resampling is the dominant factor behind improved discretization robustness. This suggests that future work should focus on architectures that combine the efficiency of Fourier-based methods with explicit control of internal representations. Hybrid models that retain spectral convolution while enforcing a stable internal discretization may provide a better balance between accuracy and robustness than standard FNO.

Additional spectral methods could also have strengthened the comparison on uniform-grid problems. A recent example is the Walsh--Hadamard Neural Operator (WHNO) \cite{whno}, which is better suited to discontinuities and shocks than standard Fourier-based operators. In particular, WHNO--FNO ensembles are shown to outperform both individual models on datasets containing sharp discontinuities. This could be especially relevant for problems such as Darcy flow, where the binary permeability field introduces strong material interfaces. Unfortunately, this work was published after much of the FNO experimentation in this thesis had already been completed and was therefore not included in the experimental comparison.

Similarly, improved anti-aliasing strategies and more accurate filtering schemes could help preserve the benefits of the UND block without unnecessarily attenuating the highest retained modes.

Overall, the central challenge remains the same: constructing neural operators whose internal representations remain faithful to the underlying continuous problem across changing discretizations. Improving this consistency, on both structured and unstructured domains, remains a key direction for making neural operators reliable tools for scientific computing and reactor modeling.