\chapter{\label{cha:method}Experimental Setup}

This chapter describes the datasets, evaluation metrics, and training
procedures used throughout the thesis. It also clarifies how models
with different output-resolution behavior are compared in the
discretization transfer experiments.

\section{Datasets}

This section should describe all datasets used in the thesis and
separate public benchmarks from the Navier-Stokes dataset generated for
this work.

\subsection{Public datasets}

List the benchmark datasets, summarize their governing equations and
data format, and explain why they are suitable for evaluating neural
operators.

\subsection{Generated Navier-Stokes dataset}

Describe how the dataset was generated, which parameters were varied,
and any quirks that later help explain model behavior. If this becomes
too detailed, the generation procedure can be moved to an appendix while
retaining a concise summary here.

\section{Evaluation Metrics}

Define the metrics used to compare models, such as the $L_2$ error and
relative error, and explain how they are aggregated over the test set.

\section{Training Details}

Summarize the optimization settings, learning-rate schedule, early
stopping or patience settings, number of epochs, and any preprocessing
steps shared across experiments.

\section{Resolution Transfer Protocol}

This section should explain how discretization invariance is evaluated
in practice.

\subsection{Train-test resolution split}

Explain that models are trained on one resolution and evaluated on other
resolutions to measure generalization across discretizations.

\subsection{Architectural comparison details}

Clarify that FNO naturally predicts on the input resolution, whereas CNO
predicts on its internal resolution, and explain how outputs are made
comparable for a fair evaluation.

